01_pca

### 2019-04-25 analysis/01_limma_all.Rmd
minor word change
now with normalized counts

---
title: "01_limma_voom"
output: md_document
---

@@

```{r}
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)
library(cowplot)
library(ggplot2)

@@

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

@@

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
```

@@

levels(colData$sex)
col.sex <- c("#5ab4ac","#d8b365")[colData$sex]
plotMDS(parentalobject,col=col.sex)
legend("bottomright",fill=c("#5ab4ac","#d8b365"),legend=levels(colData$sex))
title("Colored by Sex")
```
```{r plotMDS-colorshape}
colData$sexTissue <- paste(colData$sex, colData$tissue, sep = "_")

@@

# volcano plots

```{r volcanoplots}
# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots

lrt <- glmLRT(fit,coef=2)
topTags(lrt)

tt <- topTags(lrt,n=10000)$table

ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
```

>>>>

---
title: "01_limma_all"
output: md_document
---

@@

```{r}
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)
library(cowplot)

@@

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
```

@@

# find and print data
names(parentalobject)
head(countData)
head(parentalobject$counts)
head(parentalobject$pseudo.counts)

write.csv(parentalobject$pseudo.counts, "../results/01_pseudo.counts.csv")

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
tr$table
```

@@

levels(colData$sex)
col.sex <- c("#5ab4ac","#d8b365")[colData$sex]
plotMDS(parentalobject,col=col.sex)
legend("bottomright",fill=c("#5ab4ac","#d8b365"),legend=levels(colData$sex))
title("Colored by Sex")
```

```{r plotMDS-colorshape}
colData$sexTissue <- paste(colData$sex, colData$tissue, sep = "_")

@@

### 2019-04-25 analysis/01_pca_all.Rmd
less output
now with normalized counts

```{r setup}
library(tidyverse)
library(ggfortify) # for PCA analysis
library(cluster)
library(cowplot)

theme_rmh <- function(){ 
    theme_bw(base_size=10) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

@@

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)

head(samples)
```

```{r pca}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

@@

autoplot(pca, data = samples, colour = 'tissue') 
autoplot(pca, data = samples, colour = 'sex') 
autoplot(pca, data = samples, colour = 'treatment') 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex") 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = samples, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(counts),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')

head(pca)
```

```{r pca-es}
p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = "PCA plot in English") +
  scale_colour_discrete(name = "colour", labels = c("pink", "green", "blue")) +
  scale_shape_discrete(name = "shape", labels = c("circle", "square")) +
  theme_rmh()

es <- p + labs(subtitle = "PCA plot en español") +
  scale_colour_discrete(name = "color", labels = c("rosado", "verde", "azul")) +
  scale_shape_discrete(name = "forma", labels = c("circulo", "triangulo")) +
  theme_rmh()

enes <- plot_grid(en,es)

enes

ggsave("../figures/espanol/pca-en-es.png", width=6, height=3, dpi=300)

pdf(file="../figures/espanol/pca-en-es.pdf", width=6, height=3)
plot(enes)    
dev.off()
```

```{r subset}
for (eachgroup in levels(samples$treatment)){

@@

  countData <- counts %>% dplyr::select(one_of(savecols))

@@

  p <- autoplot(pca, data = colData, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh()

plot(en) 
}
```

>>>>

```{r setup}
library(tidyverse)
library(ggfortify) # for PCA analysis
library(cluster)
library(cowplot)
library(stringr)
library(gridExtra)
library(grid)

theme_rmh <- function(){ 
    theme_bw(base_size=14) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

treatmentcolors <-  c("control" = "#bdbdbd", 
                              "bldg" =  "#d95f02", 
                              "lay"  = "#a1d99b", 
                              "inc.d3" = "#74c476", 
                              "inc.d9"  = "#41ab5d", 
                              "inc.d17" = "#238b45", 
                              "hatch"  = "#807dba", 
                              "n5"  = "#6a51a3", 
                               "n9" = "#4a1486")

@@

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)

samples$treatment <- factor(samples$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
samples$tissue <- factor(samples$tissue, levels = 
                              c("hypothalamus", "pituitary", "gonad"))

geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
```

```{r raw}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

@@

p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")

rawPCA112 <- p + labs(subtitle = "Raw counts") +
  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
  theme_rmh() + guides(alpha = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = tissue)) 
rawPCA112

```

## with normalize edgeR data

```{r normalized}

pseudocounts <- read.csv("../results/01_pseudo.counts.csv", row.names = 1)
head(pseudocounts)
pca <- prcomp(t(pseudocounts))

p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")
p

edgRpca12 <- p + labs(subtitle = "Normalized counts from edgeR") +
  scale_alpha_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
  theme_rmh() + guides(alpha = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = tissue))
edgRpca12
```

```{r sidebyside}

sidebyside <- plot_grid(rawPCA112 + theme(legend.position = "none"),
                        edgRpca12,
                        rel_widths = c(0.4,0.6))
sidebyside

```

## subsets of normalized data show by tissue and sex

```{r tissuesex}
plotlist = list()
for (eachgroup in levels(samples$treatment)){

@@

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

@@

  p <- autoplot(pca, data = colData, alpha = 'tissue', shape = "sex", color = "treatment")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh() +
    theme_rmh()  +
  stat_ellipse(aes(linetype = sex, alpha = tissue, color = treatment)) +
   theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values = treatmentcolors) + 
  guides(color = FALSE)

legend <- get_legend(en)

en <- en  + theme(legend.position = "none")

plotlist[[eachgroup]] = en

}

main <- grid.arrange(grobs=plotlist,ncol=3)

withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
withlegend

```

## sex and timepoint within a tissue

```{r timepointsex}
plotlist = list()
for (eachtissue in levels(samples$tissue)){

  print(eachtissue)

  colData <- samples %>%
      dplyr::filter(tissue == eachtissue) %>%
      droplevels()
  row.names(colData) <- colData$V1

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

  p <- autoplot(pca, data = colData, colour = 'treatment', shape = "sex")

en <- p + labs(subtitle = eachtissue) +
  theme_rmh() +  guides(colour = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = treatment)) +
  theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values =  treatmentcolors)

legend <- get_legend(en)

en <- en  + theme(legend.position = "none")

plotlist[[eachtissue]] = en

#plot(en) 
}

main <- grid.arrange(grobs=plotlist,ncol=3)

withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
withlegend

```
